---
title: "Disengagement Analysis"
output:
  word_document: default
  html_document: default
date: "2023-12-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("~/Desktop/Research/DyadicRL/DyadicRL-Follow-up/real_data_preprocessing")
```


```{r}
library(readxl)
library(ggplot2)
library(dplyr)
```

## Definition of Disengagement

We say that a participant becomes disengaged, if we receive no observation (missing) on all of the variables in at least $P\%$ of the remaining study.

```{r}
aggregate = function(data0){
  n = dim(data0)[1]
  
  heart_data = cbind(data0[2:n,"morningHEART"], data0[1:(n-1),"dayHEART"], data0[1:(n-1),"nightHEART"])
  heart = apply(heart_data, 1, mean, na.rm = TRUE)
  heart[is.nan(heart)]=NA
  
  mood_data = cbind(data0[1:(n-1),"morningMOOD"], data0[1:(n-1),"dayMOOD"], data0[1:(n-1),"nightMOOD"])
  mood = apply(mood_data, 1, mean, na.rm = TRUE)
  mood[is.nan(mood)]=NA
  
  sleep_data = cbind(data0[2:n,"morningSLEEP"], data0[1:(n-1),"nightSLEEP"])
  sleep = apply(sleep_data, 1, sum, na.rm = TRUE)
  sleep[is.nan(sleep)]=NA
  sleep[sleep==0]=NA
  
  step_data = cbind(data0[2:n,"morningSTEPS"], data0[1:(n-1),"daySTEPS"], data0[1:(n-1),"nightSTEPS"])
  step = apply(step_data, 1, sum, na.rm = TRUE)
  step[is.nan(step)]=NA
  step[step==0]=NA
  
  new_data = data.frame(HEART = heart, MOOD = mood, SLEEP = sleep, STEPS = step)
  return (new_data)
}
```

```{r}

pairs = read_excel("BMT_Roadmap_with_Peds_2023_04_21.xlsx")

file_location = "summary_stats_removing0"
impute_file_location = "imputed_data"
residual_file_location = "residuals_heter"
missing_count = function(a){sum(is.na(a))/length(a)}

length_thres = 7*15

n = 0

list_patient = list()
list_caregiver = list()

list_raw_data = list()

for (study_ID in c(201,203)){
  study_directory = sprintf("%s/%s/",file_location,study_ID)
  caregivers = list.files(path = study_directory)
  for (caregiver_number in (1:length(caregivers))){
    compliance = FALSE
    caregiver = substr(caregivers[caregiver_number],1,8)
    if (length(which(pairs[,3]==caregiver)) > 0){
        ## can find a match
        patient = toString(pairs[which(pairs[,3]==caregiver),2])
        filename_caregiver = sprintf("%s/%s/%s.Rda",file_location,study_ID, caregiver)
        load(filename_caregiver)
        data_caregiver0 = data_one_user
        data_caregiver = aggregate(data_caregiver0)
        filename_patient = sprintf("%s/%s/%s.Rda",file_location,202, patient)
        if (file.exists(filename_patient)){
          
          load(filename_patient)
          data_patient0 = data_one_user
          data_patient = aggregate(data_patient0)
          
          if(dim(data_patient)[1] >= length_thres & dim(data_caregiver)[1] >= length_thres){
            n = n + 1
          }
          else{
            next
          }
          
          list_raw_data[[n]] = data_patient
          
          days = length_thres
          for(d in 1:days){
            data_patient[d, ] = apply(data_patient[d+1:days, ], 2, missing_count)
          }
          
          for(d in 1:days){
            data_caregiver[d, ] = apply(data_caregiver[d+1:days, ], 2, missing_count)
          }
          
          list_patient[[n]] = data_patient[1:days, ]
          list_caregiver[[n]] = data_caregiver[1:days, ]
        }
    }
  }
}
print(n)
```

There are 158 matched dyads.

We calculate the average remaining missing rate at each day. For a specific participant $i$, at the day $t$ we calculate the missing rate in the remaining days (out of 105 days) for each variable, which we denote as $\text{miss}_{i, t, j}$, where $j$ is the index for variables in MOOD, STEPs, SLEEP and HEART. Then we take average over all the participants for the groups of patients and caregivers.

```{r, out.width="50%", out.height="50%"}
library(ggplot2)
ave_patient = list_patient[[1]]
for(i in 2:n){
  ave_patient = ave_patient + list_patient[[i]]
}
ave_patient = ave_patient/n
ave_patient$Day = seq(1, length_thres)
ave_patient$Ave = (ave_patient$HEART + ave_patient$MOOD + ave_patient$SLEEP + ave_patient$STEPS) / 4

ave_caregiver = list_caregiver[[1]]
for(i in 2:n){
  ave_caregiver = ave_caregiver + list_caregiver[[i]]
}
ave_caregiver = ave_caregiver/n
ave_caregiver$Day = seq(1, length_thres)
ave_caregiver$Ave = (ave_caregiver$HEART + ave_caregiver$MOOD + ave_caregiver$SLEEP + ave_caregiver$STEPS) / 4

# convert to long table
library(tidyr)
long_patient = gather(ave_patient, key = "Variable", value = "Value", -Day)
long_caregiver = gather(ave_caregiver, key = "Variable", value = "Value", -Day)

g1 = ggplot(data = long_patient) + geom_line(aes(color = Variable, x = Day, y = Value)) + ylab("Ave Missing Rate Across Participant") + ggtitle("Patient") + theme_minimal()
g2 = ggplot(data = long_caregiver) + geom_line(aes(color = Variable, x = Day, y = Value)) + ylab("Ave Missing Rate Across Participant") + ggtitle("Caregiver") + theme_minimal()
g1
g2
ggsave("Suvival_analysis/missing_rate_patient.pdf", g1, width = 15, height = 10, units = "cm")
ggsave("Suvival_analysis/missing_rate_caregiver.pdf", g2, width = 15, height = 10, units = "cm")

```

It is noted that on day 75 there are on average around 70% missing rate in the remaining days. We set $P = 70%$ and study the drop out rate at the different days.
```{r, out.width="50%", out.height="50%"}
library(ggplot2)

data_list = list()
data_list[[1]] = list_patient
data_list[[2]] = list_caregiver

drop_list = list()
p = 0.7

for(j in 1:2){
  # dat = data_list[[j]]
  drop_list[[j]] = list()
  for(i in 1:n){
    dat = data_list[[j]][[i]]
    for(var in 1:4){
      for(d in 1:length_thres){
        if(dat[d, var] > p){
          dat[d:length_thres, var] = 0
          break
        }
        else{
          dat[d, var] = 1
        }
      }
    }
    drop_list[[j]][[i]] = dat
  }
}

ave_patient = drop_list[[1]][[1]]
for(i in 2:n){
  ave_patient = ave_patient + drop_list[[1]][[i]]
}
ave_patient = ave_patient/n
ave_patient$Day = seq(1, length_thres)
ave_patient$Ave = (ave_patient$HEART + ave_patient$MOOD + ave_patient$SLEEP + ave_patient$STEPS) / 4

ave_caregiver = drop_list[[2]][[1]]
for(i in 2:n){
  ave_caregiver = ave_caregiver + drop_list[[2]][[i]]
}
ave_caregiver = ave_caregiver/n
ave_caregiver$Day = seq(1, length_thres)
ave_caregiver$Ave = (ave_caregiver$HEART + ave_caregiver$MOOD + ave_caregiver$SLEEP + ave_caregiver$STEPS) / 4

# convert to long table
library(tidyr)
long_patient = gather(ave_patient, key = "Variable", value = "Value", -Day)
long_caregiver = gather(ave_caregiver, key = "Variable", value = "Value", -Day)

g1 = ggplot(data = long_patient) + geom_line(aes(color = Variable, x = Day, y = 1-Value)) + theme_minimal() + ylab("Ave Dropout Rate Across Participant") + ggtitle("Patient")
g2 = ggplot(data = long_caregiver) + geom_line(aes(color = Variable, x = Day, y = 1-Value)) + theme_minimal() + ylab("Ave Dropout Rate Across Participant") + ggtitle("Caregiver")
g1
g2
ggsave("Suvival_analysis/dropout_rate_patient.pdf", g1, width = 15, height = 10, units = "cm")
ggsave("Suvival_analysis/dropout_rate_caregiver.pdf", g2, width = 15, height = 10, units = "cm")

```

```{r, out.width="50%", out.height="50%"}

library(survival)
g_list = list()
g_hazard_list = list()
for(j in 1:2){
  dat = drop_list[[j]][[1]]
  dat$ALL = as.numeric(dat$HEART | dat$MOOD | dat$SLEEP | dat$STEPS)
  dat$Day = seq(1, length_thres)
  
  for(i in 2:n){
    tmp = drop_list[[j]][[i]]
    tmp$ALL = as.numeric(tmp$HEART | tmp$MOOD | tmp$SLEEP | tmp$STEPS)
    tmp$Day = seq(1, length_thres)
    dat = rbind(dat, tmp)
  }
  
  var = c("HEART", "MOOD", "SLEEP", "STEPS", "ALL")
  surv_list = list()
  hazard_list = list()
  for(v in var){
    surv_model = survfit(Surv(dat$Day, dat[, v]) ~ 1)
    # Extract the cumulative hazard function from the survival model
    cumulative_hazard <- -log(surv_model$surv)
    # Calculate the hazard function
    hazard_function <- diff(cumulative_hazard)
    surv_list[[v]] = surv_model$surv
    hazard_list[[v]] = hazard_function
  }
  surv <- data.frame(do.call(cbind, surv_list))
  surv$Day = seq(1, length_thres)
  
  hazard <- data.frame(do.call(cbind, hazard_list))
  hazard$Day = seq(2, length_thres)
  
  library(tidyr)
  long_surv = gather(surv, key = "Variable", value = "Value", -Day)
  
  g = ggplot(data = long_surv) + geom_line(aes(color = Variable, x = Day, y = Value)) + theme_minimal() + ylab("Survival Rate from Cox Model")
  if(j == 1){
    title = "Patient"
  }
  else{
    title = "Caregiver"
  }
  g = g + ggtitle(title)
  g_list[[j]] = g
  
  long_hazard = gather(hazard, key = "Variable", value = "Value", -Day)
  
  g = ggplot(data = long_hazard) + geom_line(aes(color = Variable, x = Day, y = Value)) + theme_minimal() + ylab("Hazard Rate from Cox Model")
  if(j == 1){
    title = "Patient"
  }
  else{
    title = "Caregiver"
  }
  g = g + ggtitle(title)
  g_hazard_list[[j]] = g
}
g_list[[1]]
g_list[[2]]
ggsave("Suvival_analysis/survival_rate_patient.pdf", g_list[[1]], width = 15, height = 10, units = "cm")
ggsave("Suvival_analysis/survival_rate_caregiver.pdf", g_list[[2]], width = 15, height = 10, units = "cm")

ggsave("Suvival_analysis/hazard_rate_patient.pdf", g_hazard_list[[1]], width = 15, height = 10, units = "cm")
ggsave("Suvival_analysis/hazard_rate_caregiver.pdf", g_hazard_list[[2]], width = 15, height = 10, units = "cm")
```


### Susan's suggestion

We use a different definition of disengagement. We say that an user becomes disengages, if they answer $x$ or fewer EMAs in the past 7 days. 

```{r}
dis_threshold = 7
```

```{r}
library(readxl)
pairs = read_excel("BMT_Roadmap_with_Peds_2023_04_21.xlsx")

file_location = "summary_stats_removing0"
impute_file_location = "imputed_data"
residual_file_location = "residuals_heter"

missing_count = function(a){sum(is.na(a))}

length_thres = 7*15

n = 0

list_patient = list()
list_caregiver = list()

list_raw_data = list()

for (study_ID in c(201,203)){
  study_directory = sprintf("%s/%s/",file_location,study_ID)
  caregivers = list.files(path = study_directory)
  for (caregiver_number in (1:length(caregivers))){
    compliance = FALSE
    caregiver = substr(caregivers[caregiver_number],1,8)
    if (length(which(pairs[,3]==caregiver)) > 0){
        ## can find a match
        patient = toString(pairs[which(pairs[,3]==caregiver),2])
        filename_caregiver = sprintf("%s/%s/%s.Rda",file_location,study_ID, caregiver)
        load(filename_caregiver)
        data_caregiver0 = data_one_user
        data_caregiver = aggregate(data_caregiver0)
        filename_patient = sprintf("%s/%s/%s.Rda",file_location,202, patient)
        if (file.exists(filename_patient)){
          
          load(filename_patient)
          data_patient0 = data_one_user
          data_patient = aggregate(data_patient0)
          
          if(dim(data_patient)[1] >= length_thres & dim(data_caregiver)[1] >= length_thres){
            n = n + 1
          }
          else{
            next
          }
          
          list_raw_data[[n]] = data_patient
          
          days = length_thres
          data_patient$disengage = 0
          data_caregiver$disengage = 0
          for(d in 7:days){
            data_patient$disengage[d] = as.numeric(missing_count(data_patient$MOOD[max(d-6, 1):d]) >= dis_threshold)
            if(data_patient$disengage[d] == 1){
              data_patient$disengage[d:days] = 1
              break
            }
          }
          for(d in 7:days){
            data_caregiver$disengage[d] = as.numeric(missing_count(data_caregiver$MOOD[max(d-6, 1):d]) >= dis_threshold)
            if(data_caregiver$disengage[d] == 1){
              data_caregiver$disengage[d:days] = 1
              break
            }
          }
          
          list_patient[[n]] = data_patient[1:days, ]
          list_caregiver[[n]] = data_caregiver[1:days, ]
        }
    }
  }
}
print(n)
```

```{r}
drop_list = list()
drop_list[[1]] = list_patient
drop_list[[2]] = list_caregiver

for(j in 1:2){
  dat = drop_list[[j]][[1]]
  dat$Day = seq(1, length_thres)
  
  for(i in 2:n){
    tmp = drop_list[[j]][[i]]
    tmp$Day = seq(1, length_thres)
    dat = rbind(dat, tmp)
  }
  
  library(tidyr)
  g = ggplot(data = long_surv) + geom_line(aes(color = Variable, x = Day, y = Value)) + theme_minimal() + ylab("Survival Rate")
  if(j == 1){
    title = "Patient"
  }
  else{
    title = "Caregiver"
  }
  g = g + ggtitle(title)
  g_list[[j]] = g
  
  long_hazard = gather(hazard, key = "Variable", value = "Value", -Day)
  
  g = ggplot(data = long_hazard) + geom_line(aes(color = Variable, x = Day, y = Value)) + theme_minimal() + ylab("Hazard Rate from Cox Model")
  if(j == 1){
    title = "Patient"
  }
  else{
    title = "Caregiver"
  }
  g = g + ggtitle(title)
  g_hazard_list[[j]] = g
}
g_list[[1]]
g_list[[2]]
#ggsave("Suvival_analysis/survival_rate_patient.pdf", g_list[[1]], width = 15, height = 10, units = "cm")
#ggsave("Suvival_analysis/survival_rate_caregiver.pdf", g_list[[2]], width = 15, height = 10, units = "cm")

#ggsave("Suvival_analysis/hazard_rate_patient.pdf", g_hazard_list[[1]], width = 15, height = 10, units = "cm")
#ggsave("Suvival_analysis/hazard_rate_caregiver.pdf", g_hazard_list[[2]], width = 15, height = 10, units = "cm")
```

```{r}
library(survival)
library(survey)
library(car)
library(foreign)

drop_list = list()
drop_list[[1]] = list_patient
drop_list[[2]] = list_caregiver
g_list = list()
g_hazard_list = list()
g_cum_hazard_list = list()
for(j in 1:2){
  dat = drop_list[[j]][[1]]
  dat$Day = seq(1, length_thres)
  dat$ID = 1
  for(i in 2:n){
    tmp = drop_list[[j]][[i]]
    tmp$Day = seq(1, length_thres)
    tmp$ID = i
    dat = rbind(dat, tmp)
  }
  dat$Day = as.factor(dat[, 'Day'])
  
  dat_by_day = dat %>%
  group_by(Day) %>%
  summarise(event = sum(disengage),
            total = n()) %>%
  mutate(surv = event/total)
  
  dat_by_day$hazard = c(0, diff(dat_by_day$surv))
  
  plot = dat_by_day %>%
  ggplot(aes(x = as.numeric(Day), y = surv)) +
  geom_line() +
  theme_minimal() + ylab("Disengagement Rate") + xlab("Days")
  g_list[[j]] = plot
  
  dat$Day = as.numeric(dat$Day)
  dat$disengage = as.numeric(dat$disengage)
  # model = coxph(Surv(Day, disengage) ~ 1, data=dat)
  model = survfit(Surv(Day, disengage) ~ 1, data=dat)
  # hazard = data.frame(diff(basehaz(model, centered = FALSE)$hazard))
  hazard = data.frame(model$n.event/model$n.risk)
  colnames(hazard)[1] = "Hazard"
  hazard$Day = seq(1, length_thres)
  g = ggplot(data = hazard) + geom_line(aes(x = Day, y = Hazard)) + theme_minimal() + 
        ylab("Hazard Rate from Cox Model") + xlab("Days")
  if(j == 1){
    title = "Patient"
  }
  else{
    title = "Caregiver"
  }
  g = g + ggtitle(title)
  g_hazard_list[[j]] = g
  
  # cumulative
  hazard = data.frame(cumsum(model$n.event/model$n.risk))
  colnames(hazard)[1] = "Hazard"
  hazard$Day = seq(1, length_thres)
  
  g = ggplot(data = hazard) + geom_line(aes(x = Day, y = Hazard)) + theme_minimal() + 
        ylab("Cumulative Hazard Rate from Cox Model") + xlab("Days")
  if(j == 1){
    title = "Patient"
  }
  else{
    title = "Caregiver"
  }
  g = g + ggtitle(title)
  g_cum_hazard_list[[j]] = g
}
g_list[[1]]
g_list[[2]]

g_hazard_list[[1]]
g_hazard_list[[2]]

g_cum_hazard_list[[1]]
g_cum_hazard_list[[2]]
ggsave("Suvival_analysis/survival_rate_patient.pdf", g_list[[1]], width = 15, height = 10, units = "cm")
ggsave("Suvival_analysis/survival_rate_caregiver.pdf", g_list[[2]], width = 15, height = 10, units = "cm")

ggsave("Suvival_analysis/hazard_rate_patient.pdf", g_hazard_list[[1]], width = 15, height = 10, units = "cm")
ggsave("Suvival_analysis/hazard_rate_caregiver.pdf", g_hazard_list[[2]], width = 15, height = 10, units = "cm")
```

## How many participants come back after disengagement?

```{r}
library(readxl)
pairs = read_excel("BMT_Roadmap_with_Peds_2023_04_21.xlsx")

file_location = "summary_stats_removing0"
impute_file_location = "imputed_data"
residual_file_location = "residuals_heter"

missing_count = function(a){sum(is.na(a))}

length_thres = 7*15

n = 0

list_patient = list()
list_caregiver = list()

list_raw_data = list()

for (study_ID in c(201,203)){
  study_directory = sprintf("%s/%s/",file_location,study_ID)
  caregivers = list.files(path = study_directory)
  for (caregiver_number in (1:length(caregivers))){
    compliance = FALSE
    caregiver = substr(caregivers[caregiver_number],1,8)
    if (length(which(pairs[,3]==caregiver)) > 0){
        ## can find a match
        patient = toString(pairs[which(pairs[,3]==caregiver),2])
        filename_caregiver = sprintf("%s/%s/%s.Rda",file_location,study_ID, caregiver)
        load(filename_caregiver)
        data_caregiver0 = data_one_user
        data_caregiver = aggregate(data_caregiver0)
        filename_patient = sprintf("%s/%s/%s.Rda",file_location,202, patient)
        if (file.exists(filename_patient)){
          
          load(filename_patient)
          data_patient0 = data_one_user
          data_patient = aggregate(data_patient0)
          
          if(dim(data_patient)[1] >= length_thres & dim(data_caregiver)[1] >= length_thres){
            n = n + 1
          }
          else{
            next
          }
          
          list_raw_data[[n]] = data_patient
          list_patient[[n]] = data_patient[1:days, ]
          list_caregiver[[n]] = data_caregiver[1:days, ]
        }
    }
  }
}
print(n)
```

```{r}
dis_threshold = 7
window = 7
come_back = 0
disengage = 0
missing_count = function(a){sum(is.na(a))}
for(i in 1:n){
  dat = list_raw_data[[i]]
  days = length_thres
  dat$disengage = 0
  for(d in 1:days){
    if(dat$disengage[d] == 1){
      if(d-2*window+1 >= 1 & (missing_count(dat$MOOD[max(d-2*window+1, 1):d]) <= 5)){
        come_back = come_back + 1
        break
      }
    }
    if((dat$disengage[d] == 0) & (missing_count(dat$MOOD[max(d-window, 1):d]) >= dis_threshold)){
      dat$disengage[d:days] = 1
      disengage = disengage + 1
    }
  }
}
print(come_back)
print(disengage)
```








